\documentclass[../main.tex]{subfiles} 

\renewcommand{\imageSrc}{../images/}
\renewcommand{\codeSrc}{../code/}

\begin{document}
\chapter{Real-Time Operating Systems}

\section{Scheduling}
The scheduling problem in a real-time operating system is finding a suitable order of task  execution, more specifically one that satisfies the real-time constraints. When dealing with \textit{hard real-time systems}, these constraints need to be satisfied, no matter what. The problem is twofold:

\begin{enumerate}
	\item \textbf{Are there schedules that satisfy the constraints?} \\
	Can deadlines be maintained? What's the worst case processing time? To answer these questions processing time must be bound, resources must be available and interaction between processes possible. 
	\item \textbf{If such schedule(s) exist, find one!}
\end{enumerate}

\subsection{Classic Real-Time Scheduling}
\subsubsection{Static Scheduling}
With static scheduling all decision are made \textbf{before the system is started}. Schedulability analysis is performed statically without the system running. Usually the program will consist of a loop that performs all tasks in sequence. Sometimes tasks are only executed a limited amount of times each iteration. The loop runs either continuously or is started every clock tick. The last approached is used by most programmable logic controllers.
\\\\
Interrupts are often disabled, since they are not really necessary when we know on beforehand what's going to happen. If there are interrupt routines, they are taken into account as high priority periodical tasks which run at the highest possible repetition rate of the interrupt.
\\\\
Static scheduling is only suitable for simple and immutable systems. However, system that do not change do not exist, so when anything changes the scheduling analysis needs to be repeated. This can be a problem, for example when a component needs urgent replacement by a non-identical one.

\subsubsection{Dynamic Scheduling}
Dynamic scheduling is used when static scheduling is not appropriate. 
\begin{enumerate}
	\item Perform a rough static scheduling analysis to evaluate the \textbf{worst-case processing time}. Also verify that, unless the scheduler makes a gross fault, the scheduler will be able to satisfy all timing constraints. 
	\item Let a dynamic scheduler chose the order of execution of tasks.
	\item Unless it can be proven that the scheduler always makes the right choice, test should cover as many situations as possible in order to verify timing constraints are met.
\end{enumerate}

Every mutli-tasking operating system (real-time or not) includes a scheduler. The classical qualities expected from the scheduler in an ordinary operating system consist of :
\begin{itemize}
	\item \textbf{Fairness}: All tasks should receive a fair chair of computing time.
	\item Optimization of the mean \textbf{response time} 
	\item Optimization of \textbf{available processing time} 
\end{itemize}
The real-time scheduler on the other hand has only one goal: guarantee that all real-time tasks will be finished before their deadline. 
\\\\
Independent of whether the scheduler is real-time, they can be divided in two classes:
\begin{itemize}
	\item \textbf{Non preemptive}: when the a task is allowed to run, the scheduler waits for the task to finish before allowing another task to run.
	\item \textbf{Preemptive}: the scheduler may suspend the execution of a running task when it judges appropriate.  
\end{itemize}

\begin{blockquote}
At frist sight, a preemptive scheduler who can suspend tasks in favor of more urgent tasks will obtain better results. This is not always the case because resource conflicts can arise. If  the lower priority task has locked a resource which is needed by the more urgent task, this tasks may not be suspend before this resource is freed. If this is not the case, the higher priority task would simply block, waiting for the resource.
\end{blockquote}

There are 3 techniques available for selecting the next task to run on a non-periodic real-time system.
\begin{itemize}
	\item \textbf{Fixed Priorities} \\
	Each task is assigned a fixed priority which has to be chosen with care. The priority should incorporate the urgency as well as the criticality of the task. The scheduler always selects the task with all its resource available and the highest priority. This technique is easy to implement but if there's always a task with a high priority, the lower priority task might be overrun.
	
	\item \textbf{Closest Deadline First} \\
	The scheduler selects the task closest to its deadline. This implies the deadline of each task is known. This can only be known at run time. 	
	
	\item \textbf{Least Laxity First} \\
	This scheduling technique selects the tasks with the lowest latency. If the task is started after that time, it cannot respect its deadline. The deadline and maximum execution times of tasks must be known.	
	\begin{center}
		$	\textit{laxety} = t_{\textit{deadline}} - t_{\textit{now}} - \textit{task execution time}$
	\end{center}
	\end{itemize}
	
\subsubsection{Priority Inheritance}
A scheduler always selects one of the tasks that are ready to run, which means all its resources are available. This can be a problem when several tasks need the same resource and one tasks needs the resource without it being released between subsequent runs.

\begin{exmp}
A task could have to send a message over the network to another machine. As long as the message is not entirely transmitted, no other task may send over the network. If not, what would travel on the network would be the beginning of the first message, followed by the second message, followed by the end of the first message. The resulting traffic would not make any sense.
\end{exmp}

To avoid the problem mentioned, a task can use a \textbf{mutex} on the resource he needs. If the mutex is free, the task can reserve it for itself, otherwise it waits for it to be released. The task is then suspended and only restarted when the resource is free.
\\\\
Consider the following problem called the \textbf{priority inversion} problem illustrated below.
\begin{exmp}
There are 3 tasks $T_i$ close to their deadlines with priority $i$. If there were no resource problems the scheduler would select them in the following order: $T_1, T_2, T_3$. Task $T_1$ and $T_3$ need the same resource. Assume that $T_3$ is running and using the resource and the scheduler learns that $T_1$ and $T_2$ must be run urgently. This is what would happen:
\begin{enumerate}
	\item The scheduler suspends $T_3$ which is holding the resource.
	\item Since $T_1$ cannot run, $T_2$ is chosen.
	\item After $T_2$ only $T_3$ is available and the scheduler consequently activates it.
	\item Finally $T_1$ is run as soon as $T_3$ finishes.
\end{enumerate} 
Because of the resource,$T_2$ has been executes before T1 instead of the opposite: This is called the priority inversion problem. What should have happened is that $T_3$ was allowed to finish so that the resource was released and $T_1$ could run before $T_2$.
\end{exmp}
To enforce correct behaviour, \textbf{priority inheritance} is used. The list of resources needed by each task is known by the scheduler. When a lower priority task is running and uses a necessary recource, that task temporarily inherits the more urgent task's priority untill the resource is freed.
\\\\
Notes:
\begin{itemize}
	\item The higher the number of tasks with different priorities who need the same resource, the more often this problem arises. 
	\item The longer a task uses a resource, the more this task will slow down a higher priority task even with our solution.
	\item If several tasks need a same set of resources at the same time, they must lock all resources atomically. If not, a deadlock might occur.
\end{itemize}

\subsubsection{Effect of Interrupts}
An interrupt management system behaves like a preemptive scheduler with fixed priorities. It shares the responsibility of allocating the computer's processing time with the scheduler but is hierarchically of greater importance.
\\\\
The scheduler of the operating system can only allocate the processing time left by the interrupt routines. If we aim to have the system mainly managed by the scheduler, interrupt routines should be very short. The interrupt should inform the scheduler but the latter will have to decide for itself to activate it or not.
\\\\
For timing calculations, one can try to estimate the maximum amount of time lost by interrupt routines and add this value to the task's own processing time.

\subsubsection{Periodic Tasks}
An interesting special case is that of systems that only  have periodic tasks but with a different value of periodicity. If all tasks are independent and if it is possible to satisfy all timing constraints then the scheduler that always activates the task with the highest repetition rate will meet all the deadlines. This approach is called \textbf{rate monotonic scheduling}. Unfortunately, situations where all tasks are independent are rare.

\subsection{Reservation Scheduling}
Rather than executing the scheduling algorithm every time a new task must be selected, \textbf{Reservation Scheduling} will make this choice on beforehand, as soon as the necessary information is available. 

If it is possible to schedule a task between the current time and the moment there will be no known tasks left to execute, a schedule is prepared. If the previous is not possible, the scheduler decides what tasks should be sacrificed, based on their importance (criticality) rather than their urgency. Sacrificed tasks may be replaced by shorter ones aimed at limiting the impact of sacrificing the task. Tasks that are urgent but not critical may simply be dropped.

\begin{exmp}
The considered system is an airplane. An example of a shorter task to replace the sacrificed task is the case where, if the plane cannot be saved, he is ejected. An example of a dropped task is fuel optimization when it must be leveled urgently.
\end{exmp}

This scheduling approach is implemented in two parts:
\begin{enumerate}
	\item \textbf{Pre-Scheduler} prepares a scenario.
	\item \textbf{Dispatcher} activates tasks according to the scenario. It resembles the static scheduler, it has nothing to decide.
\end{enumerate} 

\subsubsection{Reservation Scheduling Types}

There are two types of reservation schedulers for both the scheduler and the dispatcher. We're considering the scheduler first.\\\\
The first variant has the pre-scheduler simply check that the timing constraints can be satisficed and accepts or rejects tasks. The dispatcher works like a classical dynamic scheduler. This introduces the risk that the fact that there is one schedule that meets all timing requirements does not mean that all possible schedule do. The dispatcher could still make the wrong choice.
\\\\
The second variant has the pre-scheduler build itself a schedule that it has proved to meet the requirements. The dispatcher then executes it without changing anything. This variant is considered to be the best.
\\\\
The dispatcher can be preemptive or non preemptive. In the first case The pre-scheduler must decide when the dispatcher will suspend each running task. The problem here is that there must be a way to make sure a long task is not cut at the wrong moment. The solution is making use of \textbf{critical sections}. The scheduler then knows not to cut the task at that time. 
\\\\
When non preemptive scheduling is used no task can have a shorter guaranteed response time than the duration of the longest task.
\\\\
In conclusion there are 3 acceptable solutions:
\begin{itemize}
	\item \textbf{Non preemptive with short tasks}. It's up to the programmer to decide where to cut.
	\item \textbf{Preemptive with a scheduler that knows were not to cut}. The programmer must indicate these sections.
	\item \textbf{Preemptive with short critical sections}. Sections are so short that the chance of them being cut is negligible. 
\end{itemize}

\subsubsection{Assessment of Reservation Schedulers}
Reservation schedulers take into account both urgency and criticality of a task. A classic dynamic scheduler will always select the most urgent task, but this is not always the best choice. Reservation schedulers make their decision earlier so they can balance their decision based on all known factors. They can decide to sacrifice a an urgent task for a more important one.
\\\\
Why don't all real-time systems use reservation schedulers? This is because if the system has the necessary resources to never miss a deadline, a classic scheduler is able to do the job. However, since more and more real-time systems are computer controlled, overdimensioning the controlling computers is not always a good idea, especially not when they are battery operated. Reservation schedulers may in that case be the better choice.

\subsection{Example: MicroC/OS-II}
\textit{Information presented in this section is based on the book of Jean Labrosse, MicroC/OS-II, The Real-Time Kernel, Second Edition, ISBN-13:978-1-57820-103-7. The presentation is different.}
\\\\
MicroC/OS-II is a commercial product of Micrium. However, higher education institutions may use, without charge for teaching purposes, as well the sources of this software as the executable programs themselves.
\\\\
In MicroC/OS-II, each task has a different priority (a value between 4 and 60, 4 being the highest and 60 the lowest) and the scheduler is a simple dispatcher that will always launch the highest priority task. This may look rather primitive, but it is actually rather versatile because the priority of tasks may be changed dynamically. MicroC/OS-II uses that feature internally to avoid priority inversion, but it could also be used to give a task the role of pre-scheduler. However, using the same value to identify a task and his priority can be a nuisance to handle if the priority changes relatively often.
\\\\
The operating system is simply linked with the application: the main() function of the program is the OS; the code of application tasks is implemented as C functions called by this main().

\lstinputlisting[style=cstyle, label=microC/OS-II]{\codeSrc/microcosii.c}

The private memory areas of each task mus be declared as a global static variable, arrays are used to store the stack of tasks. Variables shared by several tasks (there is none in this example) must also be declared as global static variables. The EVENT structures associated with events that tasks can be waiting for must be declared as global static variables too:
\begin{itemize}
	\item Interrupts
	\item Mutex and semaphores,
	\item Task synchronization
\end{itemize}
The \texttt{main()} function initializes the operating system abd the data structures used and one or several tasks. Each task can launch other tasks. The last operation of \texttt{main()} is to start the scheduler. Initializing a task means associate with that task:
\begin{itemize}
	\item The function can execute the tasks.
	\item A pointer to the data structure of the task (where static variables are stored)
	\item A pointer to the top of the task's stack
	\item The initial priority of the task
\end{itemize}
All this is performed by the function \texttt{OSTaskCreate()}. The rest of the program consists of the functions corresponding to the tasks and of other functions.




\section{Peripherals and IO}
\section{Communication}
\section{Time Management and Synchronization}



\end{document}